---
title: "Linear Regression in Matrix Notation -- starter code"
output: html_notebook
---

This notebook explores linear regression from scratch on a tiny data set

```{r}
X = c(49, 69, 89, 99, 109)
y = c(124, 95, 71, 45, 18)
```


We can see that linear regression will fit this data fairly well if we plot the data with an abline. 

```{r}
plot(X, y)
abline(lm(y~X))
```


First, here is the solution for simple linear regression in algebraic notation:

$$ \hat{b} = \bar{y} - \hat{w}\bar{x}$$
$$ \hat{w} = \frac{\sum_{i=1}^{n} (x_i - \bar{x})(y_i - \bar{y}))}{\sum_{i=1}^{n} (x_i - \bar{x})^2}$$
Find w-hat, the estimated parameter. 

```{r}
x_mean = mean(X)
y_mean = mean(y)

w_hat = sum((X - x_mean) * (y - y_mean)) / sum((X - x_mean) ^ 2)


print(paste("w hat = ", w_hat))
```

Find b and print it. 

```{r}

b = y_mean - w_hat * x_mean

print(paste("b = ", b))
```

Compare to lm()

```{r}

plot(X, y)
abline(lm(y~X))
abline(y~w_hat * x + b)
```

### Matrix notation

$$\hat{w} = (X^TX)^{-1}X^Ty  $$


Set up matrices. Matrix notation uses a column of 1s to represent b, or $w_0$

```{r}
X = matrix(cbind(1, X), ncol=2)
X
```

First, perform 

$$X^TX$$
X transpose will be a 2x5 matrix, X is 5x2. Therefore the result of the matrix multiplication will be a 2x2  matrix.

```{r}
result <- t(X) %*% X
result
```

Find the inverse. The solve() function is built into R.

```{r}
xt_x_inv <- solve(result)
xt_x_inv
```
We could use the following formula since it is a 2x2 matrix.

$$ if A = \begin{bmatrix} a & b \\ c & d \end{bmatrix} $$
Then A inverse is:

$$ \frac{1}{ad - bc} \begin{bmatrix} d & -b \\ -c & a \end{bmatrix} $$


```{r}
matrix2 <- matrix(c(result[2,2], -result[1,2], -result[2,1], result[1,1]), nrow=2)
const <- 1 / (result[1,1]*result[2,2] - result[1,2]*result[2,1]) 
inv2 <- const * matrix2
inv2
```

Now compute the second term.

X transpose is a 2x5 matrix, y is 5x1, so the product will be 2x1.

```{r}
# your code here

term2 <- t(X) %*% y
term2
```

Now multiply the two terms. 

```{r}
final <- inv2 %*% term2
final
```

The same result is achieved from either notation. Note that the inverse was simple to compute here. Normally, computing the inverse is computationally complex, so R uses alternate methods to find the inverse. 



